---
title: "lab_02"
author: "derek willis"
date: "2024-08-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries and establish settings

```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse. If you have not installed the tidyverse already, remove the # from the next line and run it first.  
# install.packages('tidyverse')
library(tidyverse)
library(lubridate)
```

## Load Data

For this lab we'll be working with a slice of Maryland voter registration data: records of people who have registration dates this year through Aug. 7. Let's load it.

```{r}
md_voters_2024 <- read_rds("data/lab_02.rds")
```

## Examine Data

Take a look at the data, getting familiar with what it contains. Most of the columns should be self-explanatory, but there are two date columns: STATE_REGISTRATION_DATE and COUNTY_REGISTRATION_DATE. The former represents the date the person registered as a Maryland voter; the latter is when they registered in their current county. They may be the same, but if a voter moves from one county to another they will differ. This is important because if the two dates are the same, that's a pretty good indication that this is a *new* voter, someone who has no history of registering in Maryland. You'll need to keep that definition in mind.

## Answer questions

**Q1** How many voters appear to be *new* registrants since July 21, when President Joe Biden announced he would leave the race, based on the information above? Create a dataframe to hold the only those voters. Which counties have seen the most? Take a look at county population figures here and compare the two in your answer: https://worldpopulationreview.com/us-counties/maryland.

**A1** The counties with the most registrations were Montgomery and Prince George's County. These two counties have some of the highest populations in Maryland which would explain why they have more new registrations than other counties. (Used Chatgpt for help with the pipelines)

```{r}
data <- md_voters_2024 %>%
  mutate(new_voter = STATE_REGISTRATION_DATE == COUNTY_REGISTRATION_DATE)
date_of_interest <- as.Date("2023-07-21")

new_voters_since_july_21 <- data %>%
  filter(new_voter == TRUE, STATE_REGISTRATION_DATE > date_of_interest)

new_voters_df <- new_voters_since_july_21

head(new_voters_df)


county_counts <- new_voters_df %>%
  group_by(COUNTY) %>%
  summarise(count = n()) %>%
  arrange(desc(count))


print(county_counts)


```

**Q2** A better question is whether the new voters you found in Q1 are an outlier compared to other weeks this year. Using the original dataframe with 2024 registrants, use `mutate()` and lubridate to produce week numbers from the COUNTY_REGISTRATION_DATE (Lubridate has a function that turns a date into a week number. Use that to answer this question). Save that to the original dataframe. Then calculate how many voters are in each week. Describe the distribution of voters groups in a sentence - are the weeks from July 21-August 7 different from the rest, and if so, how?

**A2** The data does not seem to be an outlier if I am analyzing the data correctly because it seems to follow a similar trend as the other weeks. But I am bit confused about 

```{r}

voter_data_2024 <- md_voters_2024 %>%
  filter(year(COUNTY_REGISTRATION_DATE) == 2024)


voter_data_2024 <- voter_data_2024 %>%
  mutate(week_number = week(COUNTY_REGISTRATION_DATE))

weekly_voter_count <- voter_data_2024 %>%
  group_by(week_number) %>%
  summarise(voter_count = n()) %>%
  arrange(week_number)

start_week <- week(as.Date("2024-07-21"))
end_week <- week(as.Date("2024-08-07"))

voters_july_august <- weekly_voter_count %>%
  filter(week_number >= start_week & week_number <= end_week)

summary_stats <- weekly_voter_count %>%
  summarise(mean_voter_count = mean(voter_count),
            median_voter_count = median(voter_count),
            max_voter_count = max(voter_count),
            min_voter_count = min(voter_count))


print(weekly_voter_count)
print(voters_july_august)
print(summary_stats)

```

**Q3** Using lubridate, show which month saw the largest number of registrations. Does the answer help shed more light on Q2?
**A3** The month with the most voter registrations seems to be July with 26,284 people registering. I suppose this would support that July was an outlier in the data.

```{r}
voter_data <- md_voters_2024 %>%
  mutate(month = month(COUNTY_REGISTRATION_DATE, label = TRUE)) 
monthly_voter_count <- voter_data %>%
  group_by(month) %>%
  summarise(voter_count = n()) %>%
  arrange(desc(voter_count))


max_month <- monthly_voter_count %>%
  filter(voter_count == max(voter_count))


print(monthly_voter_count)
print(max_month)
```

**Q4** Maryland has, officially, five regions: https://www.visitmaryland.org/article/maryland-regions. Modify the original dataframe again to add a `region` column based on the `COUNTY` value using case_when. Are you going to have to do this for every Maryland county plus Baltimore City? Yes, you are. Which regions and parties account for the most new voters?

**A4** 

```{r}
voter_data <- md_voters_2024 %>%
  mutate(region = case_when(
    COUNTY %in% c("Allegany", "Garrett", "Washington") ~ "Western Maryland",
    COUNTY %in% c("Carroll", "Frederick", "Howard", "Montgomery") ~ "Capital Region",
    COUNTY %in% c("Anne Arundel", "Baltimore", "Baltimore City", "Harford", "Cecil") ~ "Central Maryland",
    COUNTY %in% c("Calvert", "Charles", "Prince George's", "St. Mary's") ~ "Southern Maryland",
    COUNTY %in% c("Caroline", "Dorchester", "Kent", "Queen Anne's", "Somerset", "Talbot", "Wicomico", "Worcester") ~ "Eastern Shore",
    TRUE ~ "Unknown"  
  ))

voter_data <- voter_data %>%
  mutate(STATE_REGISTRATION_DATE = as.Date(STATE_REGISTRATION_DATE, format="%Y-%m-%d"),
         COUNTY_REGISTRATION_DATE = as.Date(COUNTY_REGISTRATION_DATE, format="%Y-%m-%d"))

new_voters <- voter_data %>%
  filter(STATE_REGISTRATION_DATE == COUNTY_REGISTRATION_DATE)

new_voters_by_region_party <- new_voters %>%
  group_by(region) %>%
  summarise(new_voter_count = n()) %>%
  arrange(desc(new_voter_count))

print(new_voters_by_region_party)

```

**Q5** What's an additional question you could ask of this data?

**A5** Why are the registrations numbers for the Eastern Shore and Western Maryland so low?
