---
title: "Data Analysis"
author: "Jack, Madison, and Josh"
date: "Nov. 9"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 

In this notebook, we are doing data analysis on the political emails.

## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}
library(tidyverse)
library(janitor)
```

## Load and Cleaning Data

We pulled the data from a database of political emails. It is a massive dataframe, and for some reason, at first, it seemed like the download is not pulling all of the data required. Had to make a change and select "stream all rows," but the number of rows still doesn't match up with the number on the website.

```{r}
# Load required data
emails <- clean_names(read_csv("data/emails.csv"))
emails


# Path to data should be loaded from folder "data" i.e. read_csv("data/name_of_data.csv")

# Clean required data and prepare for analysis if needed. 


```
```{r}
glimpse(emails)
```
There are some problems here that we are going to need to get worked out. The first is that the date format is in date/time format; I think this should work, but would need to consult.

In terms of defining rows and columns, there are 35,965 rows (huge dataset) and 14 columns. I couldn't save the data to GitHub. I had to remove the data.

It should be pretty easy to answer the questions of which candidates send the most. The figuring out of what the emails are about may be harder; it seems like it might be a little hard to parse content with how it's set up (very messy). We may have to adjust our questions to make them more quantitative than qualitative.

```{r}
emails <- emails |> 
  #TAKING OUT THE ROWID BECAUSE IT DOES NOT ADD ANYTHING TO THE DATA. SEEMS LIKE JUST AN EXTRA COLUMN THAT WE CAN REMOVE
  select(-rowid) |> 
  #renaming the date so it's explaining that it is also time.
  rename(date_and_time = date)
emails
```
-30-