---
title: "lab_03"
author: "Derek Willis"
date: "2024-08-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries and establish settings

```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse
library(tidyverse)

# Load janitor
library(janitor)
```

## Load data

```{r}
# Load MD slice of FEC House candidate contribution data
md_contribs <- read_rds("data/lab_03.rds")
```

## Fix header issues

You'll need to fix one issue with the column names using janitor. Write a code block to do that, making sure that you save the results to the same dataframe.

```{r}
md_contribs <- md_contribs |>
  clean_names()

md_contribs
```


## Answer questions

Q1: Which seven cities in Maryland had the largest *number* of contributions from donors? How many did each have? To answer this question, you'll have to clean the messy "contributor_city" field.

Note: To answer this question, you'll need to see every variation of the "contributor_city" for the top seven cities, which you should identify through some initial analysis. Then find variations of them and change those variations into the most common version.

A1: BETHESDA	1213,	CHEVY CHASE	704,	POTOMAC	490,	BALTIMORE	692, ANNAPOLIS	492,	SILVER SPRING	408

```{r}

md_contribs |> 
group_by(contributor_city) |>
summarise(count = n(),
            total_amount = sum(contribution_receipt_amount)) |>
arrange(desc(total_amount))


#md_contribs <- md_contribs %>%
  #mutate(contributor_city = case_when(
  # str_detect(contributor_city, "BALRIMORE|BLATIMORE") ~ "BALTIMORE",  
   # str_detect(contributor_city, "N BETHESDA|NORTH BETHESDA") ~ "BETHESDA",   
   # str_detect(contributor_city, "ANAPOLIS") ~ "ANNAPOLIS",
    #str_detect(contributor_city, "SILVER SPRINGS") ~ "SILVER SPRING",  
   # TRUE ~ contributor_city
  #))




  
```

Q2: Find the 5-digit ZIP code in Maryland that had the highest total amount of contributions (by dollar value). You will need to use mutate to ensure that you have a zip code column with the first five characters from the existing zip code column.

Where is that zip code and what kind of area is it, demographically? Which committee is the largest recipient of money from that zip code? How many of the top five recipients are Maryland candidates?

A2: 20815, 20817, 20854, 20814, 21401 These areas are Chevy Chase, Bethesda, Potomac, ect. which are all very rich communities. For the zipcode 20815 the committee with the largest recipient of money was Jamie Raskin for Congress.Two of the five candidates are from Maryland. 

```{r}
#md_contribs |>
 # mutate(contributor_zip = as.character(contributor_zip)) |>
    #mutate(zip = str_sub(contributor_zip, start=1L, end=5L)) ) 

#md_contribs |>
#group_by(zip) |>
  #summarise(contribution_receipt_amount = n(),
            #total_amount = sum(contribution_receipt_amount)) |>
#arrange(desc(total_amount))


largest_recipient <- md_contribs |>
  filter(zip == 20815) |> 
  group_by(committee_name) |>  
  summarise(contribution_receipt_amount = n(),
            total_amount = sum(contribution_receipt_amount)) |>
arrange(desc(total_amount))|>
  top_n(5, wt = total_amount) 

print(largest_recipient)



```

Q3: Let's look at top individual donors. Write code to generate a list of top donors using name and city, adding up the total amount of contributions. What does the original data say about the top two donors - who are they and what do they do? Do you see anything in the list of top 10 donors that would make you uncertain about the accuracy of these results?

A3: Name: KRACOV, DANIEL Occupation: ARNOLD & PORTER LLP Name: ALLEN, MARGARET	Occupation: BALTIMORE	AGM FINANCIAL SERVICES INC. The tenth person on the list is unemployed so that may be a red flag if they are a top contributor. Also the amounts seem pretty low.

```{r}
top_fr<-md_contribs|>
  group_by(contributor_name, contributor_city, contributor_employer) |> 
 summarise(contribution_receipt_amount = n(),
            total_amount = sum(contribution_receipt_amount)) |>
arrange(desc(total_amount))


top_10_donors_fr <- top_fr |> 
  head(10)

print(top_10_donors_fr)

```
