---
title: "lab_01"
author: "Derek Willis"
date: "2024-08-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Turn off scientific notation
options(scipen=999)
```

## About this lab

To complete this lab, you need to: \* run existing code as directed (look for **Task**). \* modify existing code as directed (look for **Task**). \* write code in empty codeblocks provided to answer questions included (look for **Q**). \* write out the answer in the form of a complete sentence in the space given (look for **A**). When you are finished, commit changes and push to your personal GitHub repo, then submit the URL to this document on ELMS.

Remember to add any new verbs or concepts to your reference notebook!

## Load libraries and establish settings

**Task** Run the codeblock below to load the Tidyverse. To run the codeblock, hit the little green play button in the upper right corner of the gray area (the codeblock) below.

```{r}
# Load the tidyverse. If you have not installed the tidyverse already, remove the # from the next line and run it first.  
# install.packages('tidyverse')
library(tidyverse)
```

## Load Data

We're going to load the slice of WinRed contribution data from West Virginia, which is in a file called `lab_01.rds`

**Task** Using the example from the pre_lab, write code to load the RDS file and assign it to a good variable name.

```{r}
westvirginia_winred_contributions <- read_rds("lab_01.rds")
```

## Examine Data

**Task** Run the codeblock below to use glimpse to get a sense of the column names, types and sample data.

**Q1** How many rows and columns does this dataframe have?

**A1**
There are 40,441 rows and 24 columns.
```{r}
glimpse(westvirginia_winred_contributions)
```

## Answer questions

**Q2** Which West Virginia city provided the largest number of contributions? How many?

**A2**Charleston

```{r}
westvirginia_winred_contributions |>
  group_by(city) |>
  summarise(
    count_contribs = n(),
  ) |>
  arrange(desc(count_contribs))
```

**Q3** Which West Virginia cities accounted for the fewest contributions? How many of them, roughly speaking?

**A3**There were 63 cities with only one contribution in West Virginia.

```{r}
westvirginia_winred_contributions |>
  group_by(city) |>
  summarise(
    count_contribs = n(),
  ) |>
  arrange(count_contribs)
```

**Q4** What's your best educated guess to explain why the city identified in question 2 had more contributions compared to the cities identified in question 3? Hint: you won't find the answer in the data, you'll need to do some quick web research. You do not need to write any code for this question, so no codeblock is provided.

**A4** Charelston has more people and likely more money so they would be donating significantly more than small towns. 

**Q5** What recipient got the largest total amount of money? How many contributions did they receive and what was the average donation?

**A5**

```{r}
westvirginia_winred_contributions |>
  group_by(memo_text) |>
  summarise(
    count_contribs = n(),
    total_amount = sum(amount),
    mean_amount = mean(amount),
    median_amount = median(amount)) |>
  arrange(desc(count_contribs))
```

**Q6** What is that top recipient and what explains its popularity? Hint: this will require web research, you won't be able to answer this purely from the data set. You do not need to write any code for this question, so no codeblock is provided.

**A6**The top recipiant was for Trump Natinal Committee JFC. It is the official fundraising site for Trump so people will tend to dontate to that site to avoid potential scams. Additionally, Trump is running for president so it would make sense that he would have more contributions that other republicans or other republican organizations. 

**Q7**. Using the answer from A5 as a starting point, produce a summary table that has the following information for each recipient: \* count of contributions \* total amount of all contributions \* the average contribution amount \* the median contribution amount \* the smallest contribution amount \* the largest contribution amount

Store it as a variable object called `wv_summary` so that you can examine it with the data viewer. What newsworthy nuggets jump out at you that you could imagine dropping into a news story? Write a few summary sentences with overall findings.

**A7** For one story we could go to Charleston and see why they have the most contributors and see why that area has so many contributors. We could also use find the city with the highest amounts given and go to that city to see how and why people have so much money and are willing to give it to the Trump administration. We could also look into why there were 1298	refunds which seems very high, it could have something to do with the subcrption scam by the Trump administration.

```{r}
wv_summary <- westvirginia_winred_contributions |>
  filter(memo_text != 'Refund of Earmarked Contribution') |> 
  group_by(state) |>
  summarise(
    count_contribs = n(),
    total_amount = sum(amount),
    mean_amount = mean(amount),
    median_amount = median(amount),
    min_amount = min(amount),
    max_amount = max(amount)
  ) |>
  arrange((total_amount))
```
