---
title: "lab_07"
author: "derek willis"
date: "2024-08-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## You will need

-   A Census API key

## Load libraries and establish settings

You'll need to load two packages for this: the tidyverse and tidycensus.

**Task** load these two packages

```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse.
library(tidyverse)
library(tidycensus)
library(dplyr)
```

## Setup Census API

You'll need your Census API Key:

**Task** set your API Key *only* if you haven't already

```{r echo=FALSE}
census_api_key("YOUR API KEY", overwrite=TRUE)
```
```{r echo=FALSE}
census_api_key("3d3f88866c4da579b432a35db04b8f20fe94e438", overwrite=TRUE)
```
## Load and view ACS variables

You'll also need to load the list of American Community Survey variables from the 2022 5-year estimates so you can identify the codes you need:

**Task** load the variables from the 2022 5-year ACS and save them to a dataframe

```{r}

variables_2022 <- load_variables(2022, "acs5", cache = TRUE)


df_variables <- as.data.frame(variables_2022)


head(df_variables)

```

## Answer questions

**Q1** What is the Maryland jurisdiction with the lowest median age, according to the ACS 2022 5-year estimates? You will need to find the variable for median age first.

**A1** 	Baltimore city, Maryland is the Maryland jurisdiction with the lowest median age, according to the ACS 2022 5-year estimates

```{r}

md_data <- get_acs(geography = "county", variables = "B01002_001", 
                   state = "MD", year = 2022, survey = "acs5")


lowest_median_age <- md_data[which.min(md_data$estimate), ]


lowest_median_age



```

**Q2** Which Maryland counties have a median age of at least 50? You can use the dataframe you just made to answer this, but you must write code to display only those counties with a median age of at least 50.

**A2** Talbot and Worcester County counties have a median age of at least 50.

```{r}

md_counties_age_50 <- md_data[md_data$estimate >= 50, ]


md_counties_age_50
```

## Ticket-Splitting

Ticket-splitting (voting for candidates from multiple parties) in today's political environment seems like a rarity, but it does happen, even in states like Maryland. For the remaining questions you'll be looking at cast vote records from Frederick County - Maryland's most competitive county - for the 2022 general election. You'll be trying to identify the zip codes where ticket-splitting is happening the most and to learn more about those places. To do that, you'll need to read in two CSV files, join them and then also bring in data from the Census API.

### Setup

Load the two CSV files in the `data` folder: one containing the ballot choices of Frederick voters in five races (governor, comptroller, attorney general and U.S. House) and the other containing information about Frederick's voting locations. You'll need to join them together.

```{r}

ballots <- read.csv("data/frederick_cvr.csv")
locations <- read.csv("data/frederick_precincts.csv")

joined_data <- dplyr::inner_join(ballots, locations, by = "precinct")


head(joined_data)
```

**Q3** Let's answer some basic questions about ticket-splitting: how many voters voted for different parties in the governor and comptroller's race? How many voted for the same party in both races? What percentage of all voters split their tickets in those races? In how many cases did voters choose the same party in all five contests? What percentage of all voters does that represent?

**A3** split_ticket_count : 12081, same_party_count: 95431, split_ticket_percentage: 11.25888, same_party_all_five
0, same_party_percentage: 0 (not sure why its running but the values are 0)



```{r}
# Count the number of voters who voted for different parties in the governor and comptroller's race
split_tickets <- joined_data[joined_data$governor_lt_governor != joined_data$comptroller, ]
split_ticket_count <- nrow(split_tickets)

# Count the number of voters who voted for the same party in both races
same_party_count <- nrow(joined_data[joined_data$governor_lt_governor == joined_data$comptroller, ])

# Calculate the percentage of voters who split their tickets
split_ticket_percentage <- (split_ticket_count / nrow(joined_data)) * 100

# Count the number of voters who voted for the same party in all five contests
same_party_all_five <- nrow(joined_data[joined_data$governor_lt_governor == joined_data$comptroller &
                                        joined_data$governor_lt_governor == joined_data$attorney_general &
                                        joined_data$governor_lt_governor== joined_data$us_house, ])

# Calculate the percentage of voters who voted for the same party in all five contests
same_party_percentage <- (same_party_all_five / nrow(joined_data)) * 100

# Display the results
list(
  split_ticket_count = split_ticket_count,
  same_party_count = same_party_count,
  split_ticket_percentage = split_ticket_percentage,
  same_party_all_five = same_party_all_five,
  same_party_percentage = same_party_percentage
)

```

**Q4** Now let's look for the zip codes where voters preferred Republicans in every race *except* the governor's race, where Dan Cox lost to Wes Moore by a wide margin. Write code to generate a dataframe of zip codes with the number of those voters for each of those zip codes, ordering the dataframe so that the zip code with the most ticket-splitters appears first.

**A4** The zipcodes with the most Cox splitters were 21702, 21701, 21703

```{r}
non_cox_voters <- joined_data %>%
  filter(governor_lt_governor == "DEM" & 
         comptroller == "REP" &
         attorney_general == "REP" & 
         !is.na(zipcode)) 


zip_code_splitters <- non_cox_voters %>%
  group_by(zipcode) %>%
  summarise(splitter_count = n()) %>%
  arrange(desc(splitter_count))


print(zip_code_splitters)


```

**Q5** Let's find out more about those zip codes from A4. Using tidycensus, get a dataframe of Maryland zip codes and their median ages, then join it to the dataframe you produced in A4. NOTE: you'll need to ensure that the two columns you use for the join are the same datatype. Summarize your findings; how would you describe the zip codes with more non-Cox Republicans compared to those with fewer? Where are those leading zip codes located?

**A5** The average age for the Cox splitters in the top zipcodes have an average age of around 35. zipodes 21702, 21703, and 21701: These zipcodes are located in Frederick County, Maryland. Frederick is a suburban area with significant population growth, which may explain more moderate Republicans, as suburban voters tend to differ from both urban and rural voters.

Median Age: You can now describe whether ZIP codes with more non-Cox Republicans have older or younger populations on average.

```{r}

maryland_zip_data <- get_acs(
  geography = "zcta",
  variables = "B01002_001", 
  year = 2020,
  survey = "acs5"
) %>%
rename(zip_code = GEOID, median_age = estimate) %>%
select(zip_code, median_age)
```
```{r}

maryland_zip_codes <- c("21702", "21703", "21701")  


maryland_zip_data <- maryland_zip_data %>%
  filter(zip_code %in% maryland_zip_codes)
```

```{r}

maryland_zip_data <- maryland_zip_data %>%
  rename(zipcode = `zip_code`)


non_cox_voters <- non_cox_voters %>%
  rename(zipcode = zipcode) 

```

```{r}
maryland_zip_data$zipcode <- as.character(maryland_zip_data$zipcode)
non_cox_voters$zipcode <- as.character(non_cox_voters$zipcode)

```



```{r}
merged_df <- non_cox_voters %>%
  left_join(maryland_zip_data, by = "zipcode")

```


```{r}

non_cox_voters <- joined_data %>%
  filter(governor_lt_governor == "DEM" & 
         comptroller == "REP" &
         attorney_general == "REP" & 
         !is.na(zipcode))


non_cox_voter_counts <- non_cox_voters %>%
  group_by(zipcode) %>%
  summarize(total_non_cox_republicans = n()) 


non_cox_voter_counts$zipcode <- as.character(non_cox_voter_counts$zipcode)
merged_df$zipcode <- as.character(merged_df$zipcode)


#merged_df <- merged_df %>%
 # left_join(non_cox_voter_counts, by = "zipcode") %>%
  #mutate(total_non_cox_republicans = ifelse(is.na(total_non_cox_republicans), 0, total_non_cox_republicans)) # Fill NA values with 0


summary_df <- merged_df %>%
  group_by(zipcode) %>%
  summarize(
    total_non_cox_republicans = sum(total_non_cox_republicans, na.rm = TRUE),
    median_age = median(median_age, na.rm = TRUE)
  ) %>%
  arrange(desc(total_non_cox_republicans))

head(summary_df)
```